The Quantum Internet --- Towards the Singularity

### 2017-09-10 Tim Byrnes/rmp.tex
Begin merging Tim Byrnes space stufff

\title{The Quantum Internet | Towards the Singularity}

\author{Peter Rohde}
\affiliation{UTS}

\author{Other dudes}

\affiliation{Other affiliations}

\author{Tim Byrnes}
\affiliation{State Key Laboratory of Precision Spectroscopy, School of Physical and Material Sciences,
East China Normal University, Shanghai 200062, China}
\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 3663 Zhongshan Road North, Shanghai 200062, China}
\affiliation{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}
\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China} 
\affiliation{Department of Physics, New York University, New York, NY 10003, USA}

\author{Jonathan P. Dowling}
\affiliation{Hearne Institute for Theoretical Physics, Department of Physics \& Astronomy, Louisiana State University, Baton Rouge, Louisiana 70803-4001, USA}

\date{\today{}}

\begin{abstract}
[Abstract]
\end{abstract}

%\pacs{81.05.Uw,68.37.-d,73.20-r}

@@

\tableofcontents{}

\section{Foreword}
\label{foreword}

[Previous sections]

\section{Space based quantum communication}
\label{sec:space}

[TB: This introductory paragraph probably needs integrating with what is written already, such as adding section references]

One of the major hurdles that must be overcome before quantum networks to gain widespread commercial use is to span large distances between nodes.  The key technological and commercial centers of the world are spread worldwide, hence connections over intercontinental distances are necessary, with nodes separated by several thousands of kilometers. This distance is one of the major challenges facing the creation of a large-scale quantum quantum networks. The most convenient way of transmitting quantum information is using optical fibers. However, it is well-known that due to the exponential scaling of loss with distance, the upper limit is of the order of several hundred kilometers. This shortcoming has led to intense research relating to extending these distances using quantum repeaters. In a quantum repeater, one creates long-distance entanglement by joining together entangled pairs, established over multiple short-distance links, through entanglement swapping \cite{sangouard11}. Much research has been done in this direction, but such quantum repeaters remain a challenging task primarily due to the lack of a reliable quantum memory -- a medium that can store quantum information and convert it reliably between photons \cite{lvovsky2009optical,simon2010quantum,bussieres2013prospective, heshami2016quantum}.

An alternative to optic fiber communication is using terrestrial free-space links. Particular wavelength ranges have low absorption, and it is possible to propagate photons across large distances. Impressive experimental demonstrations of teleportation with free-space entanglement by Zeilinger and co-workers \cite{ursin07,ma2012quantum,yin2013lower} over 100km have been achieved. However, extending to longer distances have been problematic.  Photon loss in free space links is affected by weather conditions and other effects, such as pollution, that degrade visibility. The Earth’s curvature provides an absolute upper limit, depending on the elevation of the source and receiver. For example, for an observer on a 30m tower, the horizon is at a distance of 20km, a relatively short distance. The observatory used in the experiments performed by the experiments of Ref. \cite{ursin07,ma2012quantum} were are at an elevation of 2393m, which allowed for the 143km transmission.

One attractive possibility to overcome these issues is using space-based quantum communication. In this scenario, satellites orbiting the Earth would act as nodes of the quantum network, which could store, send and receive quantum information. At altitudes where satellites orbit the Earth, photon loss due to scattering is negligible, and photons can propagate across extremely long distances. The main loss in this case is caused by diffraction, due to the finite diameter of the receiver and transmitter. For example, using reasonable diameters it is in the region of 40-80 dB loss for low Earth orbit \cite{aspelmeyer2003long,liao2016ground}. Involving ground-based quantum network nodes is also not an inherent problem as 80\% of the atmosphere by mass is located within the first 12km in altitude. Ground-to-satellite and satellite-to-ground quantum communication has already been demonstrated, as will be discussed more in the next section. Such a satellite-based quantum communication system is naturally suited to various tasks, in particular quantum cryptography which does not require quantum memories. But with the addition of quantum memories the capabilities of such a quantum network could be greatly enhanced, making many of the applications listed above a possibility at the global scale.

Naturally, such a space-based quantum network comes with enormous challenges that much be overcome before it is implemented. Creating even a short-distance quantum network is currently technologically challenging, let alone one that is loaded onto a satellite transmitting photons across distances potentially of the order of the diameter of the Earth. The cooperation of space agencies is necessary even for initial experiments. As we describe more in the next section, several recent positive results has made the technology an immediate possibility in realizing such a global level quantum network. In this section, we summarize the current technological state of the art internationally, and discuss some of the remaining major challenges.

\begin{figure}
\includegraphics[width=\columnwidth]{figspace1}
\caption{Various possibilities for space-based quantum communication. (a) Satellite-to-satellite quantum communication \cite{byrnes2017lorentz}; (b) Ground-to-satellite quantum communication \cite{armengol08}. }
\label{figspace1}
\end{figure}

\subsection{International efforts for space-based quantum communication}

\begin{figure*}
\includegraphics[width=2\columnwidth]{figspace4}
\caption{The Chinese Micius quantum communications satellite.  (a) Schematic of the satellite and ground stations used to observe the entangled photons \cite{popkin17}.  [TB: Took image from Science News, may need permission] (b) Attenuation during entanglement distribution from Ref. \cite{yin2017satellite}.  (c) Fidelities achieved for teleportation of various states as marked from Ref. \cite{ren2017ground}.  
}
\label{figspace4}
\end{figure*}

\subsubsection{China: QUESS}

In 2016, the group at the University of Science and Technology led by Jian-Wei Pan launched QUESS (Quantum Science Experiment Satellite), a quantum communication satellite \cite{gibney16,xin11}. The main feature of the satellite is an ultrabright source of polarization entangled photons at 810 nm, generated by parametric down conversion.  The source is capable of emitting $ 5.9 \times 10^6 $ photon pairs per second with a fidelity of $ \sim 0.91 $.  The other important piece of technology is the acquiring, pointing, and tracking technology, which allows the ground stations to track the location of the satellite as it moves across the sky, and vice versa.  This is achieved using lasers at separate wavelengths, and allows the photon transmission and the ground-based receivers to be pointing to each other to within $ \sim 1.2 $ mrad \cite{yin2017satellite}.  The major aims of the project are to perform QKD between Xinglong and Urumqi (a distance of 2500 km); test Bell’s inequality across a distance of 1200 km; and perform teleportation between the satellite and Ali in Tibet (see Fig. \ref{figspace4}).  The final aim is to perform QKD between Beijing and Vienna (a distance of 7500 km).

At the time of writing three main results have been reported.  The first is an entanglement distribution experiment where a Bell violation was observed between ground stations at Delingha and Lijiang (separated by 1203 km) and Delingha and Urumqi (separated by 1120 km) \cite{yin2017satellite}.  The satellite is in an orbit at a elevation of $ \sim 500$ km, and is in view from the observatories for a duration of $ \sim 275 $ s.
The attenuation of the photons during the downlink transmission is shown in Fig. \ref{figspace4}(b). Taking into account of the distances of the photon transmission, the attenuation rates are far better than the best performance of optical fibers at 0.16 dB/km \cite{yin2013lower} and even theoretical loss limits.  The Bell violation that was recorded in this experiment was at the level of $ 2.37 \pm 0.09 > 2 $, and the nonlocality of the entanglement was confirmed.

The second experiment performed quantum teleportation from the ground to the satellite \cite{ren2017ground}. In the experiment, a single polarization encoded photon defines the qubit to be sent, and the satellite acts as the receiver.  The entangled photons are generated on the ground, at the observatory in Ngari, Tibet. As with the first experiment above, the satellite is in orbit and sweeps across the sky during which the teleportation must be achieved.  The longest distance where the teleportation was successful was at a distance of $ \sim 1400 $ km when the satellite emerges from the horizon.  When directly overhead, the satellite is at a distance of $ \sim 500 $ km.  As Fig.  \ref{figspace4}(c) shows, the fidelity of the types of states teleported are all above the theoretical classical bound of $2/3$, with an average of $ 0.80 \pm 0.01 $.

A third experiment performed quantum key distribution in a satellite-to-ground configuration, where the ground station was at Xinglong \cite{liao2017satellite}. The protocol that was used is the decoy-state BB84 protocol, which is robust against a photon number splitting attack.  A key rate of between 1 to 12 kbit/s was achieved while the satellite was visible, depending upon the location in the sky (see Fig. \ref{figspace4}(d)).

\subsubsection{Japan: SOCRATES}

SOCRATES (Space Optical Communications Research Advanced Technology Satellite) is a microsatellite developed by NICT (National Institute of Information and Communications Technology) \cite{horiuchi2015view,toyoshima2015current,takenaka2017}. This is a $ 48 $ kg satellite with volume $ 50^3 $ cm$^3$ that demonstrates multi-purpose (i.e. both classical and quantum) optical communications. Its main mission is to demonstrate laser communication in space. As one of its subgoals, the on-board equipment is adapted to perform QKD experiments. The satellite is equipped with photon source where the polarization can be changed in non-orthogonal directions.  
Initial results indicated polarization preservation of photons emitted from the satellite, which is a preliminary step towards performing protocols such as BB84 \cite{carrasco2016leo}. More recent results showed quantum limited  satellite-to-ground communication was possible, attaining a quantum bit error rate below 5 \% \cite{takenaka2017} during the closest approach of the satellite to the ground at a distance of $ 744 $ km.  The downlink was established up to distances of over 1000 km, although the error rate was higher for longer distances.

\subsubsection{Singapore: Cubesat}

The group at the National University of Singapore also has taken the approach using nanosatellites, with the capability of generating correlated photon pairs on-board \cite{tang2016generation}.  
The nanosatellite is 1.65 kg and has the components necessary for creating and detecting parametric down converted photon pairs. The nanosatellite approach greatly reduces the cost for launching into space, and the launch itself was performed by the India’s space agency. As the generation and detection are both on the same satellite, no quantum communication channel (to other satellites or Earth) was established in the experiment, although it showed the space worthiness of the basic quantum optical components.

\subsubsection{Europe: Retroreflector satellite}

European groups used laser-ranging satellites fitted with corner cube retro-reflectors to demonstrate quantum communication through the atmosphere \cite{villoresi08,vallone15}. A ground based photon source sent up photons towards the satellite, and reflected down again to an Earth based observatory. This showed the ability of photons to travel through the atmosphere into space and back, with a bit error ratio at the level of 5\%. Although no optical sources or detectors were placed on the satellite, there has been long standing active interest in a European space-based quantum communication project, led in particular by the group of Anton Zeilinger at the University of Vienna \cite{armengol08}.  The aims of such projects are to perform space-based QKD, with terrestrial free-space experiments \cite{ursin07,ma2012quantum} as part of the overall effort.

\subsubsection{Canada: QEYSSat}

The QEYSSat (Quantum EncrYption and Science Satellite) proposes a microsatellite that incorporates a quantum receiver \cite{jennewein2014qeyssat}. The satellite acts as a trusted node, where photons would be transmitted from Earth-based sources. This could be used to perform QKD between two locations on Earth that are separated by large distances. No photon sources or quantum memories would be included on the QEYSSat.

\begin{figure}
\includegraphics[width=\columnwidth]{figspace2}
\caption{Satellites employing quantum technologies from various groups across the world: Japan \cite{horiuchi2015view}, Italy \cite{vallone15}, Canada \cite{jennewein2014qeyssat}, Singapore \cite{tang2016generation}, and China \cite{gibney16}.}
\label{figspace2}
\end{figure}

@@

\subsection{Future challenges}

\subsubsection{Coverage}

Currently there are only two satellites with capability of sending and/or receiving photons through space, the Chinese QUESS and the Japanese SOCRATES satellites.  In order to achieve world-wide coverage multiple satellites would be necessary such that there would be a constellation of satellites that are visible at all times, anywhere on Earth, in a similar way to GPS satellites. In the case of the QUESS satellite, the orbit is such that it passes over the observatories once per day, with a visibility window of $ \sim 275 $ s.  The coverage of a low Earth orbit satellite of elevation 500 km has a maximum radius of $ \sim 2400 $ km from the point where the satellite is directly overhead due to the curvature of the Earth.  In practice it will be less than this due to obstacles and due to atmospheric influence, which reduce it to distances of the order of $ \sim 1000 $ km (see Fig. \ref{figspace4}). By launching satellites at higher orbits, it will be possible to increase the coverage, which will lead to extra challenges in establishing a satellite link.  Some of the technologies that will need to be developed are larger-size telescopes, better tracking systems, and wave-front correction through adaptive optics \cite{liao2017satellite}.  Currently the communications are performed at night due to the laser wavelengths that are used, but plans for using telecommunication wavelengths give the possibility of daytime quantum communication.

From the point of view of security, we anticipate that the initial quantum networks for quantum key distribution will be based on trusted nodes.  Due to the difficulty of tampering of space-based trusted nodes, it is likely that this will offer a sufficient level of security for most applications. Thus using a constellation of satellites employing quantum key distribution can achieve long-distance cryptography by simply relaying the information, either by multiple satellite-to-ground or a satellite-to-satellite configuration.  For the ultimate security without the use of trusted nodes, one would employ alternative protocols that do not require line-of-sight quantum communication, such as the E91 protocol \cite{ekert1991quantum}.  As this requires entanglement distribution, storage, and purification, this would be most likely a second generation technology after the trusted node quantum key distribution network is already established.

\subsubsection{High precision applications}

For applications such as quantum clock synchronization and fundamental experiments testing physics at large length scales and velocities, it is likely that extremely high fidelities of the quantum operations will be necessary.  For example, for clock synchronization, even current atomic clocks used as national standards have an accuracy at the level of $ 10^{-14} $.  Thus such high precision experiments will almost certainly follow the less demanding quantum key distribution experiments.  As the precision of the technology improves, other sources of error may appear which may require correction.  It is well-known from existing GPS satellites, it is crucial to account for relativistic effects, due to time-dilation and the gravitational red shift. Not accounting for these effects would seriously compromise the GPS system, with inaccuracies equating to an error of the order of 10 km/day. This is due to the high velocities of LEO satellites, traveling at speeds of  $ \beta = 10^{-5} $ times the speed of light. Such relativistic effects can affect the entanglement in the presence of diffracting photons \cite{gingrich03}.  Even for single photon transmission, polarization encoded photons can give rise to correction at the order of $ \beta $ \cite{byrnes2017lorentz}.  Using relativistically invariant entanglement distribution protocols have been proposed by several authors to avoid such effects, which would otherwise require correction \cite{yurtsever02,li2003relativistic,byrnes2017lorentz}.

\section{Conclusions}
\label{conclusions}

[Conclusions]

\section{Acknowledgments}
T. B. is supported by the Shanghai Research Challenge Fund; New York University Global Seed Grants for Collaborative Research; National Natural Science Foundation of China (Grant No. 61571301); the Thousand Talents Program for Distinguished Young Scholars (Grant No. D1210036A); and the NSFC Research Fund for International Young Scientists (Grant No. 11650110425); NYU-ECNU Institute of Physics at NYU Shanghai; and the Science and Technology Commission of Shanghai Municipality (Grant No. 17ZR1443600).

J. P. D. would like to acknowledge support from the US Air Force Office of Scientific Research, the Army Research Office, the National Science Foundation, and the Northrop-Grumman Corporation.

>>>>

\title{The Quantum Internet --- Towards the Singularity}

@@

@@
